Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching

### 2025-10-01 review_responses.tex

Section IV titled ``Programs, Methodology, and Results'' is overloaded and blends experimental setup with results and interpretation. \textbf{(R1.3) For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).} \\
Conceptually, the work serves primarily as a benchmarking study rather than a novel methodological contribution. While the comparison is comprehensive, the novel losses (DTW\_Envelope, SIMSE\_Spec) are adaptations of known techniques and are not framed with sufficient theoretical justification. \textbf{(R1.4) There is also limited interpretability analysis; for example, the paper does not explore why DTW performs well on modulated noise or why SIMSE outperforms L1 in certain scenarios.} Additionally, the synthesizers used are very simple (only 2--3 parameters), and all experiments are conducted in-domain, limiting the generalizability of the findings to real-world synthesis environments with more complex parameter spaces or out-of-domain targets. \\
\textbf{(R1.5) The manuscript would also benefit from clearer practical implications.} While the authors argue that sound-matching is a creative process and reject the idea of a universal loss, they stop short of offering actionable heuristics. \textbf{Some guidance (e.g., Use DTW for AM synthesis; SIMSE for subtractive synthesis) would strengthen the paper’s relevance to practitioners.} \\
\textbf{(R1.1) The paper would be improved by formally stating additional hypotheses that are currently only implicit.} For example, the observed agreement between manual and automatic evaluation metrics, the effectiveness of the novel losses, and the usability of iterative differentiable optimization as a general strategy are all important insights, but are not positioned as testable hypotheses upfront. \\

@@

  \item The Introduction concludes with a concise statement of our contributions:

  \begin{displayquote}
      The contributions of this paper include (1) an evaluation of multiple differentiable losses across multiple synthesis methods, (2) the introduction and justification of loss functions utilizing DTW and SIMSE (3) a discussion on the utility and agreement between manual and automatic evaluation metrics (4) a nomenclature of different sound-matching approaches and unsolved issues in the field. 
  \end{displayquote}
\end{itemize}

\subsection{R1.2 Improve Background and Related Work}
\label{R1.2}
\noindent\textbf{Reviewer:}\\
\noindent\textbf{The current division of the manuscript into Introduction, Background, and Past Works results in significant conceptual overlap and repetition. Historical framing, loss function taxonomy, and method classification are spread across all three sections. A clearer organizational approach would be to merge Background, and Past Works into a unified Background and Related Work section, where formal definitions, theoretical framing, and literature review are presented coherently. This would allow the Introduction to focus solely on the motivation, problem framing, and contribution.}\\

\noindent\textbf{Response:} \\
We agree that the organization and writing in these sections could be improved. We merged \emph{Background} and \emph{Past Works} into a unified section, \textbf{II. Background and Related Work}. We applied major edits and rewrites to these sections to improve reading and remove redundancies between the various subsections while taking precautions against compromising the flow.
\textbf{II. Background and Related Work} contains the following topics in order:  
\begin{itemize}
  \item A formal definition of sound-matching and its components.
  \item Definitions of synthesizers, losses, and optimization heuristics. 
  \item A chronological review of prior sound-matching approaches.
  \item A discussion of gaps in the field and four critical issues.
  \item A justification for our methodology, and how our experiments address the two critical issues of \emph{Loss and Synthesizer Selection}.
\end{itemize}

@@

\subsection*{R1.3 Organize Methodology, Results, and Discussion.}
\label{R1.3}
\noindent\textbf{Reviewer:}\\
\textbf{Section IV titled ``Programs, Methodology, and Results'' is overloaded and blends experimental setup with results and interpretation. For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).}
\\

\noindent\textbf{Response:} \\
We agree that the earlier draft blurred the boundaries between setup, results, and discussion. In the revised manuscript we implemented the reviewer’s suggestion as follows:

\begin{itemize}
  \item \textbf{III. Experimental Setup and Results A-E} sections now exclusively describe the methodological components: the synthesizer programs, the definitions of losses, the optimization procedure, evaluation metrics, and statistical tests. Each subsection is focused on describing how the experiments are designed, without interpretation of outcomes. For instance, the synthesizer section defines parameter ranges, while the evaluation section specifies how statistical significance is tested. The only partial overlap is at the beginning of the \textbf{III-F Best Performers for Each Program}, which serves as a transition between results and discussion.

  \item Experimental analysis are presented in \textbf{III-F Best Performers for Each Program} and \textbf{III-G Consistency in Rankings}, which together form the \emph{Results} component. Section III-F reports which losses performed best for each synthesizer, while Section III-G quantifies the consistency of rankings across bootstrap resamples. Keeping all numerical findings here avoids overlap with later interpretation.

  \item \textbf{IV. Discussion} is structured into four dedicated subsections:  
    \begin{enumerate}
      \item \textbf{IV-A. Key Findings}: summarizes the central empirical trends, such as the dependency of loss performance on synthesis method.  
      \item \textbf{IV-B. Practical Recommendations}: translates results into concrete recommendations for practitioners and researchers.  
      \item \textbf{IV-C. Illustrating Loss Landscapes}: adds visualizations that explain why certain losses succeed or fail in specific contexts.  
      \item \textbf{IV-D. Applicability to More Complex Synthesizers}: discussion on how observed patterns extend beyond the simplified testbed to richer, real-world synthesizers.  
    \end{enumerate}

@@

  \item To improve clarity, we also relocated explanatory material about experiment design (e.g., parameter settings, optimizer configuration) previously scattered across the results section into Sections III-A–E, ensuring that the results sections remain strictly empirical, while the discussion is reserved for interpretation. 
\end{itemize}

Overall, this restructuring produces a clearer separation of topics: methodology is isolated in Section III-A–E, empirical findings are concentrated in Section III-F–G, and broader interpretation is reserved for Section IV. We believe this major restructuring and rewrites of appropriate sections has improved flow of the manuscript and made the boundary between methodology, results, and discussion much clearer.

\subsection*{R1.4 Justification for \DTWEnv{} and \SIMSESpec}
\label{R1.4}
\noindent\textbf{Reviewer:} \\
\textbf{\noindent While the comparison is comprehensive, the novel losses (DTW\_Envelope, SIMSE\_Spec) are adaptations of known techniques and are not framed with sufficient theoretical justification. }\\

@@

    Figure~\ref{fig:loss_landscape_noisebp} [\textbf{Figure 4 in paper}] shows the HP-Noise results. \LoneSpec{}, \SIMSESpec{}, and \JTFS{} exhibit clear minima at the correct parameter (red dashed line), whereas \DTWEnv{} remains relatively flat, making gradient-based optimization more difficult.

    Figure~\ref{fig:loss_landscape_ampmod} [\textbf{Figure 5 in paper}]  shows the S-Noise-AM results. Here, \DTWEnv{} produces the smoothest and most informative landscape, while spectrogram losses are minimized near the target but lack consistent correlation with parameter distance. \JTFS{} remains largely flat. This aligns with \DTWEnv's superior performance in amplitude-modulated synthesis observed in our experiments.
\end{displayquote}

\begin{figure*}[ht]
    \centering
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{images/experiment_plots/comparing_loss_landscapes_normalized.png}
        \caption{Loss landscapes for \BPNoise{} with only a high-pass filter parameter. 
        \LoneSpec{}, \SIMSESpec{}, and \JTFS{} show clear global minima near the correct parameter, while \DTWEnv{} remains flat around the target.}
        \label{fig:loss_landscape_noisebp}
    \end{minipage}%
    \hfill
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{images/experiment_plots/comparing_lanscapes_1_1d_normalized.png}
        \caption{Loss landscapes for a simplified \AmpMod{} synthesizer with only amplitude modulation parameter. 
        \DTWEnv{} exhibits the smoothest and most informative landscape, explaining its superior performance in amplitude-modulated synthesis.}
        \label{fig:loss_landscape_ampmod}
    \end{minipage}
\end{figure*}

While this may not be the "qualitative" or "spectral" analysis, we believe it is instructive on the mechanisms underlying our quantitative findings and highlight the importance of loss landscape navigation as an open area for future research (See also~\nameref{R4.5}).

@@

    Nonetheless, we believe that the insights here can extend naturally to more complex domains. Depending on the feature of interest, the observed dependence of loss performance on synthesis method will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.

    \DTWEnv{} measures very specific features of audio, and its use in the recommended settings---where low frequency amplitude modulation matching is required---would likely yield the desired results regardless of synthesizer complexity. Likewise, the use of \SIMSESpec{} or \LoneSpec{} for matching filter cut-offs is likely to be successful regardless of the underlying sound. 
    \end{displayquote}

\subsection{R1.7 Provide Practical Takeaways}
\label{R1.7}
\noindent\textbf{Review:} \\
\noindent \textbf{Provide practical takeaways or usage guidelines for different loss-synth combinations. } 
\\

\noindent\textbf{Response} \\
In \textbf{IV.A Key Findings}, we discuss important findings and provide answers to our research questions. The following section, \textbf{IV.B Practical Recommendations}, distills our findings into actionable heuristics for practitioners.

@@

\subsection*{R2.3 Clear Takeaways}
\noindent\textbf{Reviewer:}\\
\textbf{This work addresses the problems on ``Loss Selection'' and ``Synthesis Selection.''  However, the results as presented here are largely inconclusive (i.e. it is not imminently clear how to select losses/synthesizer pairings), leaving very little in the way of clear takeaways sparking future research.\\}

\noindent\textbf{Response:} \\
We acknowledge the reviewer’s concern that our main result may appear inconclusive. However, we see this as an important finding in itself: much of the prior literature implicitly over-claims general effectiveness of particular loss functions based on narrow or in-domain settings. Our revisions now explicitly emphasize that the lack of a “silver bullet” for loss functions is a core contribution of this work. By showing that loss effectiveness depends on the synthesis context, the inconclusiveness can be seen as evidence against over-generalization, and an argument for more thorough, context-sensitive benchmarks in measuring sound similarity.\\

As for clear takeaways, we've added two new sections which address the valid issues raised here and by Reviewer 1.

\begin{itemize}
  \item \textbf{IV.B Practical Recommendations} which distills our findings into actionable guidelines for practitioners. DTW\_Env is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation being strongly advised). We also note that our results challenge prior claims about JTFS’s universal superiority for mesostructures.
  \item \textbf{IV.D Applicability to More Complex Synthesizers} that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing. While acknowledging the limitations of our simplified testbed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings.
\end{itemize}
For a more detailed response, see our response to Reviewer 1 in Sections~\nameref{R1.6} and \nameref{R1.7}. Our response to \nameref{R1.5} may also be of interest.

@@

\subsection{\textbf{R4.1} Use of Appropriate Symbols}
\noindent\textbf{Reviewer:}\\
\textbf{The notation \textbf{[for the equation introduced in II.A]} is confusing. In the manuscript, $t$ is used to denote the target sound, but in audio/signal processing $t$ usually represents time.  
A clearer alternative would be to write $x_\theta$ and $x_0$, so that $x$ always denotes a sound waveform, with the subscript indicating whether it is generated from parameters $\theta$ or given as the target.  
This avoids overloading $t$ and improves readability.}
\\

\noindent\textbf{Response:}
This is a helpful suggestion, and we have made the switch from $x$ and $t$ to $x_{\theta}$ and $x_0$ throughout the manuscript.

\subsection{\textbf{R4.2} Optimal Targets}
\noindent\textbf{Reviewer:}\\
\textbf{You mentioned "There is no optimal or correct version of an 8-bit snare sound".
This statement is true when you are given only the text description of the sound. However, if you have the target sound $x_0$  the target sound itself is "correct", isn't it? So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?\\}

\noindent\textbf{Response:}
The original statement referred to how realistically in sound-design, a ``correct" solution rarely exists due to our subjective experience of sound and personal goals. The existence of an objective sound-matching goal is only true in \textit{in-domain replication} scenarios, which are far removed from real life applications. To reduce confusion, the section has been rephrased as follows:
\begin{displayquote}
    Moreover, the ``correct'' solution may not exist in a strict sense: depending on the artistic goal, multiple parameter settings may yield acceptable or even preferable outputs.  In the following sections we will discuss the various components of sound-matching, the subjectivity of ``correct'' solutions, and important issues in the field.
\end{displayquote}

@@

\subsection{R4.3 Hearing Tests}
\label{}
\noindent\textbf{Reviewer:}\\
\textbf{Also, in the blinded hearing test, the two listeners are the authors, it seems a bit too biased to me. Not sure how difficult is it to add one more listener, but I believe having an extra sound engineer/musician who is not one of the author would yield a more convincing listening results.
}
\\

\noindent{\textbf{Response:}}
The listening tests were done without use of any labels or previous exposure to outputs, which we believe minimizes the risk of bias. Additionally the strong correlation between the two listeners adds to our confidence in using the listening test results as the ground truth. Unfortunately, adding another listener was not possible given the allotted time for the revisions. If the reviewer feels that this is necessary, it can be done for the next round of reviews.

@@

\subsection{\textbf{R4.4} STFT Losses and Time-Shifts}
\noindent\textbf{Reviewer:}\\
\textbf{"If one of these tones is slightly shifted in time, then the spectrogram will no longer overlap". This is true if we don't apply any techniques like DTW to realign them, right?
So the STFT losses doesn't work for the cases where there's time shift?}
\\

\noindent\textbf{Response:}
What we refer to as STFT losses refers to the near ubiquitous approach of applying L1/L2 to STFT spectrograms. This approach does not account for two sounds that are nearly identical but slightly time-shifted~\cite{vahidi2023mesostructures,turian2020sorry}. We've shown in this work that DTW can be used in loss functions to account for envelope time-shifts.

\subsection{\textbf{R4.5} Can We Overfit? Is it a Concern?}
\label{R4.5}
\noindent\textbf{Reviewer:}\\
\textbf{One final question that is just out of curiosity.
In your iterative algorithm, you always run it for 200 steps, right?
In my field of machine learning, if we set 200 steps, the result in the final step might not be the optimal one due to overfitting. So sometimes we obtain the optimal results at around 100 steps, and the model gets worse after this.
Do you have the same thing in your algorithm? Or the final step is always the best step?
}
\\

\noindent\textbf{Response:} 
Since we are not training a model that will be applied to other scenarios, overfitting here just means that the parameter updates are stuck around a local or global minimum. This means that \textit{if} the target parameters have been reached, subsequent updates will not diverge from it (or at least stay very close). We share some visualizations of the gradient update paths in 2D, which we felt was not statistically robust enough for the paper. Figures~\ref{fig:p0_simse_iterations} and~\ref{fig:p2_dtw_iterations}, we see that once the gradient updates (small red dots connected by lines) take the parameters from the starting position (green circle) to the target (green star), subsequent updates fluctuate slightly and randomly around the target.

### 2025-10-10 review_responses.tex

\noindent\textbf{Response:} \\
We agree that the organization and writing in these sections could be improved. We merged \emph{Background} and \emph{Past Works} into a unified section, \textbf{II. Background and Related Work}. We applied major edits and rewrites to these sections to improve reading and remove redundancies between the various subsections while taking precautions against compromising the flow.
\textbf{II. Background and Related Work} contains the following topics in order:  
\begin{itemize}
  \item A formal definition of sound-matching and its components.
  \item Definitions of synthesizers, losses, and optimization heuristics. 
  \item A chronological review of prior sound-matching approaches.
  \item A discussion of gaps in the field and four critical issues.
  \item A justification for our methodology, and how our experiments address the two critical issues of \emph{Loss and Synthesizer Selection}.
\end{itemize}

\subsection*{R1.3 Organize Methodology, Results, and Discussion.}
\label{R1.3}
\noindent\textbf{Reviewer:}\\
\textbf{Section IV titled ``Programs, Methodology, and Results'' is overloaded and blends experimental setup with results and interpretation. For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).}
\\

\noindent\textbf{Response:} \\
We agree that the earlier draft blurred the boundaries between setup, results, and discussion. In the revised manuscript we implemented the reviewer’s suggestion as follows:

\begin{itemize}
  \item \textbf{III. Experimental Setup and Results A-E} sections now exclusively describe the methodological components: the synthesizer programs, the definitions of losses, the optimization procedure, evaluation metrics, and statistical tests. Each subsection is focused on describing how the experiments are designed, without interpretation of outcomes. For instance, the synthesizer section defines parameter ranges, while the evaluation section specifies how statistical significance is tested. The only partial overlap is at the beginning of the \textbf{III-F Best Performers for Each Program} section, which serves as a transition between results and discussion.

@@

\subsection{\textbf{R4.4} STFT Losses and Time-Shifts}
\noindent\textbf{Reviewer:}\\
\textbf{"If one of these tones is slightly shifted in time, then the spectrogram will no longer overlap". This is true if we don't apply any techniques like DTW to realign them, right?
So the STFT losses doesn't work for the cases where there's time shift?}
\\

\noindent\textbf{Response:}
What we refer to as STFT losses refers to the near ubiquitous approach of applying L1/L2 to STFT spectrograms. This approach does not account for two sounds that are nearly identical but slightly time-shifted~\cite{vahidi2023mesostructures,turian2020sorry}. We've shown in this work that DTW can be used in loss functions to account for envelope time-shifts.

\subsection{\textbf{R4.5} Can We Overfit? Is it a Concern?}
\label{R4.5}
\noindent\textbf{Reviewer:}\\
\textbf{One final question that is just out of curiosity.
In your iterative algorithm, you always run it for 200 steps, right?
In my field of machine learning, if we set 200 steps, the result in the final step might not be the optimal one due to overfitting. So sometimes we obtain the optimal results at around 100 steps, and the model gets worse after this.
Do you have the same thing in your algorithm? Or the final step is always the best step?
}
\\

\noindent\textbf{Response:} 
Since we are not training a model that will be applied to other scenarios, overfitting here just means that the parameter updates are stuck around a local or global minimum. This means that \textit{if} the target parameters have been reached, subsequent updates will not diverge from it (or at least stay very close). We share some visualizations of the gradient update paths in 2D, which we felt was not statistically robust enough for the paper. In Figures~\ref{fig:p0_simse_iterations} and~\ref{fig:p2_dtw_iterations}, we see that once the gradient updates (small red dots connected by lines) take the parameters from the starting position (green circle) to the target (green star), subsequent updates fluctuate slightly and randomly around the target.

### 2025-10-10 sections/summary_weakness_conclusion.tex
Updates from Overleaf

\textbf{Weaknesses:} 
While we cannot prove that a universally best similarity measure does not exist, we can advocate for more creativity in the field based on our findings. Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes a variety of loss functions and synthesis methods. However, there are many other methods of synthesis and sound-similarity that can be combined in practically infinite ways. Due to this large search space, we set arbitrary parameters for the various signal processing functions. We used bare-bones versions of STFT and JTFS with fixed parameters. We did not test complex synthesizers using parallel and sequential DSP functions. Arbitrary hyperparameters such as learning rate and max number of iterations were selected for the DL pipeline, and only the RMSProp optimizer was tested.

>>>>

\textbf{Weaknesses:} 
While we cannot prove that a universally best similarity measure does not exist, we can advocate for more creativity in the field. Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes a variety of loss functions and synthesis methods. However, there are many other methods of synthesis and sound-similarity that can be combined in practically infinite ways. Due to this large search space, we set arbitrary parameters for the various signal processing functions. We used bare-bones versions of STFT and JTFS with fixed parameters. We did not test complex synthesizers using parallel and sequential DSP functions. Arbitrary hyperparameters such as learning rate and max number of iterations were selected for the DL pipeline, and only the RMSProp optimizer was tested.
